---
title: "Machine Learning Project"
author: "Bill Lancaster"
date: "September 10, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =TRUE,  cache=TRUE, fig.cap=FALSE)
```

## Predict Exercise

Goal: Predict the manner if which 6 participants completing an exercise to lift a weight. The prediction has a code of A, B, C, D, or E.

There are two datasets. One is for training the model. The second file has 20 records that are to be our test subjects for our prediction.

## Preparing the data

First, I investigated the data, which had 160 columns of data with 19622 records.

After viewing the data, there were a lot of empty fields and NAs so I only imported the columns of data that would be relevant to the analysis. 

With the training data, I partitioned to the data with 70% of the records to use for the model and 30% to use to test the model before I made the prediction of the final testing records.


```{r get data}
library(caret)

rd = read.csv("~/pml-training.csv") #import raw data
dim(rd)
d <-rd[,c(8:11, 37:49, 60:68, 84:86, 102, 113:124, 140, 151:160)] #relevent variables 
dim(d)
set.seed(1500)
inTrain <- createDataPartition(y=d$classe,
                               p=0.7, list=FALSE)
train <- d[inTrain,]
test <- d[-inTrain,]
dim(train)
dim(test)
```
The model will be created using a Random Forrest with cross validation of 3 folds. The model will be created with the train data set. 

```{r model, echo=TRUE}

library(randomForest)
mrf <- randomForest(classe~ .,data=train,
             cv.fold=3, prox=TRUE)
print(mrf)
```

Now we will run it on the test data from our partitioned data.


```{r test model, echo=TRUE}
pred <- predict(mrf,test)
confusionMatrix(pred, test$classe)
```

The out of sample accuracy of the model is .9958 so the out of sample error would be expected to be approximately .42% (1-.9958 = .0042 *100). 

The final task is to make the prediction for the 20 test subjects for the final test. 

```{r model prediction, echo=TRUE}
trd = read.csv("~/pml-testing.csv") #import raw data
testing <-trd[,c(8:11, 37:49, 60:68, 84:86, 102, 113:124, 140, 151:160)]
predict(mrf,testing)
```
## Quiz Results

####Using the predictions above, scored 100% correct when automatically scored.
